On finiteness of structures produced deterministically by co-transcriptional folding

### 2018-11-14 ttt.tex
11/14 en

In this section, we prove that unary oritatami system can form infinitely at delay 1 and arity 2 deterministically and moreover that the only infinite conformations which its oritatami system can yield is only the zig-zag conformation shown in Fig.\ref{TTT_zigzag}.

@@

Let us consider the stabilization of the $i$-th bead $a_i$ upon $C_{i-1}$. The bead cannot collaborate with any succeeding bead $w[i+1],w[i+2],\cdots$ at delay 1. There are just two ways to get stabilized at delay 1. One way is to be bound to another bead. The other way is through a $tunnel\ section$. A tunnel section consists of four beads that occupy four neighbors of a point (Fig.\ref{TTT_tunnel_intro}). Assume that four of the six neighbors of a point $p$ are occupied by beads $a_{j_1},a_{j_2},a_{j_3},a_{j_4}$ while the other two are free. We call such beads as $p$ $inside\ of\ a\ tunnel$ and such beads as $p^\prime$ $entrance\ of\ a\ tunnel$ without a case that $p^\prime$ is inside of a tunnel. If the beads $w[i-2]$ and $w[i-1]$ are stabilized respectively at one of the two free neighbors and at $p$ one after another, then the next bead $w[i]$ cannot help but be stabilized at the other free neighbor. In this way, $w[i]$ can get stabilized without being bound.

If a bead is stabilized through a tunnel section, then it can provide two binging capabilities and create tunnel sections.

@@

\begin{theorem}[Tunnel Troll Theorem]
  Let $\Xi$ be an unary oritatami system of delay 1 and arity 2. If a bead is stabilized through a tunnel section, then it consume some binging capabilities.
\end{theorem}

\subsection{Proof of Tunnel Troll Theorem}
Assume $\Xi$ is deterministic. Let us represent its transcript $w$ as $w = w_1 w_2 w_3 \cdots$ for beads $w_1 , w_2 , w_3 , \cdots \in \Sigma = \{e\}$. Each of these beads is stabilized either by being bound or through a tunnel section (or by both). How they are stabilized can be described by a binary sequence $S$ of $b$'s (bound) and $t$'s (tunnel section); priority is given to $t$, that is, $S[i] = t$ if the $i$-th bead $w_i$ is stabilized not only by being bound but also through a tunnel section. 
On the other hand, each of beads are bound either inside of tunnel or outside (Fig. \ref{TTT_position1}). If a bead is stabilized at inside of tunnel, then its successor is already decided the position either inside of a tunnel or outside. Moreover, If a bead is stabilized at outside of a tunnel, then its position is an entrance of a tunnel or otherwise.

Tunnel sections have three possible shape with considering symmetry such as acute turn, straight and obtuse turn (Figure \ref{TTT_tunnel}). Let us focus entrances of a tunnel such as $p^\prime$, then Entrances have two possible shape (Figure \ref{TTT_enterpoint}).

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{lr}

      \begin{minipage}{0.2\hsize}
        \begin{tikzpicture}

@@

      \begin{minipage}{0.2\hsize}
        \begin{tikzpicture}

@@

          \begin{scope}[xshift=2cm, yshift=2cm]
            \fill(0,0) circle [radius=0.1];
            \node[above] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(60:1.2)},right] {A};
            \node[transform canvas={shift=(-60:1.2)},right] {B};

            \node[transform canvas={shift=(120:1.2)},left] {c};
            \node[transform canvas={shift=(-120:1.2)},left] {d};
          \end{scope}

          \node at (2,-0.5) {$t_0$};
        \end{tikzpicture}
      \end{minipage}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \fill(0,0) circle [radius=0.1];
            \node[below] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(180:1.2)},left] {c};
            \node[transform canvas={shift=(-120:1.2)},left] {d};
          \end{scope}

          \node at (2,-0.5) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \fill(0,0) circle [radius=0.1];
            \node[above] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(120:1.2)},left] {c};
            \node[transform canvas={shift=(180:1.2)},left] {d};
            \node[transform canvas={shift=(-120:1.2)},left] {e};
            \node[transform canvas={shift=(0:1.2)},below] {n};
          \end{scope}

          \node at (2,-0.5) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

@@

  \end{itemize}
\end{itemize}

\begin{figure}[h]
  \begin{center}
    \begin{tabular}{cc}

      \begin{minipage}{0.48\hsize}
        \begin{tikzpicture}
          \draw[thick, rounded corners = 10pt]
          (-2,2)--(5,2)--(5,-3)--(-2,-3)-- cycle;

@@

          \foreach \theta in {60,-60,120,-120,180}{
            \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.5)}](0.1,-0.1)--(-0.1,0.1);
          }

          \fill(0:1.5) circle [radius=0.1];
          \draw[->] (0:0.1)--(0:1.4);

          \node[left] at (-60:1.5) {A};

          \begin{scope}[shift=(0:3)]
            \fill(0,0) circle [radius=0.1];

            \foreach \theta in {120,60,-60}{
              \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1)--(0.1,0.1);
              \draw[transform canvas={shift=(\theta:1.5)}](0.1,-0.1)--(-0.1,0.1);
            }

            \draw[->] (180:1.4)--(180:0.1);
            \draw[->] (-120:0.1)--(-120:1.4);

            \node[below] at (180:1.5) {$w_{i-2}$};
            \node[right] at (0:0) {$w_{i-1}$};
            \node[right] at (-120:1.5) {$w_i$};

            \node[left] at (120:1.5) {B};
            \node[left] at (60:1.5) {C};
            \node[left] at (0:1.5) {D};

            \begin{scope}[shift=(-120:1.5)]
              \draw(0,0) circle [radius=0.1];
              \foreach \theta in {-120,-60}{
                \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1) rectangle (0.1,0.1);
              }

              \node[right] at (-120:1.5) {e};
              \node[right] at (-60:1.5) {f};
            \end{scope}

@@

          \node at (1.5,-3.5) {Pattern 1};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.48\hsize}
        \begin{tikzpicture}
          \draw[thick, rounded corners = 10pt]
          (-2,2)--(4.5,2)--(4.5,-4.5)--(-2,-4.5)-- cycle;

@@

          \foreach \theta in {60,0,120,-120,180}{
            \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.5)}](0.1,-0.1)--(-0.1,0.1);
          }

          \draw[->] (-60:0.1)--(-60:1.4);

          \begin{scope}[shift=(-60:1.5),shift=(0:1.5)]
            \fill(0,0) circle [radius=0.1];
            \fill(180:1.5) circle [radius=0.1];

            \foreach \theta in {120,60,0,-60}{
              \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1)--(0.1,0.1);
              \draw[transform canvas={shift=(\theta:1.5)}](0.1,-0.1)--(-0.1,0.1);
            }

            \draw[->] (180:1.4)--(180:0.1);
            \draw[->] (-120:0.1)--(-120:1.4);

            \node[below] at (180:1.5) {$w_{i-2}$};
            \node[right] at (0:0) {$w_{i-1}$};
            \node[right] at (-120:1.5) {$w_i$};

            \node[left] at (120:1.5) {A};
            \node[left] at (60:1.5) {B};
            \node[left] at (0:1.5) {C};

            \begin{scope}[shift=(-120:1.5)]
              \draw(0,0) circle [radius=0.1];
              \foreach \theta in {180,-120,-60}{
                \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1) rectangle (0.1,0.1);
              }

              \node[right] at (180:1.5) {e};
              \node[right] at (-120:1.5) {f};
              \node[right] at (-60:1.5) {g};
            \end{scope}

@@

          \node at (1.25,-5) {Pattern 2};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Case of $S[i-2..i] = tbt$}
    \label{TTT_tunnelC_enter_usingTunnel}
  \end{center}
\end{figure}

@@

      \begin{minipage}{0.48\hsize}
        \begin{tikzpicture}
          \draw[thick, rounded corners = 10pt]
          (-3,2)--(2.5,2)--(2.5,-3)--(-3,-3)-- cycle;

          \foreach \theta in {60,-60,120,0}{
            \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.5)}](0.1,-0.1)--(-0.1,0.1);
          }

          \fill(180:1.5) circle [radius=0.1];
          \fill(0:0) circle [radius=0.1];

          \draw[transform canvas={shift=(180:1.5)}] (105:1.4) edge[bend right] (105:0.1);
          \draw[->] (180:1.4) -- (180:0.1);
          \draw[->] (-120:0.1) -- (-120:1.4);

          \node[below] at (180:1.5) {$t_1$};
          \node[right] at (0:0) {$t_2$};
          \node[left] at (-120:1.5) {$t_3$};

          \node[right] at (120:1.5) {A};
          \node[right] at (60:1.5) {B};
          \node[right] at (0:1.5) {C};

          \begin{scope}[shift=(-120:1.5)]
            \draw(0,0) circle [radius=0.1];
            \foreach \theta in {-120,-60,180}{
              \draw[transform canvas={shift=(\theta:1.5)}](-0.1,-0.1) rectangle (0.1,0.1);
            }

          \node at (-0.25,-3.5) {Pattern 1};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Case of $S[i-2..i] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw[thick, rounded corners = 10pt]
      (-1.5,2)--(3,2)--(3,-2.5)--(-1.5,-2.5)-- cycle;

      \draw[transform canvas={shift=(-120:1.5)}](-0.1,-0.1)--(0.1,0.1);
      \draw[transform canvas={shift=(-120:1.5)}](0.1,-0.1)--(-0.1,0.1);

      \fill(0:0) circle [radius=0.1];
      \fill(0:1.5) circle [radius=0.1];
      \fill[shift=(0:1.5)](40:1.5) circle [radius=0.1];

      \draw[->] (-120:1.4) edge[bend left] (-120:0.1);
      \draw[->] (0:0) -- (0:1.3);
      \draw[->,transform canvas={shift=(0:1.5)}] (0:0.1) edge[bend right] (40:1.3);

      \draw [shift=(60:1.5)] (-0.1,-0.1) rectangle (0.1,0.1);
      \draw [shift=(-60:1.5)] (-0.1,-0.1) rectangle (0.1,0.1);

      \node[below] at (-120:1.5) {$v_{i-2}$};
      \node[above] at (0:0) {$v_{i-1}$};
      \node[above] at (0:1.5) {$v_i$};
      \node[left, shift=(0:1.5)] at (40:1.5) {$v_{i+1}$};

      \node[right] at (60:1.5) {A};
      \node[right] at (-60:1.5) {B};

    \end{tikzpicture}
    \caption{Case of $S[i$}
    \label{TTT_case_of_o}
  \end{center}
\end{figure}

>>>>

In this section, we consider on finiteness of structures produced deterministically at delay 1. In result, cases of arity 1 and 3 can only yield finite structures which is size of $\mathcal{O} (n)$, and cases of arity 4 and more can only yield finite structures which is size of $mathcal{O} (n^2)$, and a case of arity 2 can yield infinite structures but they are only the zig-zag conformation shown in Fig.\ref{TTT_zigzag}.

%In this section, we prove that unary oritatami system can form infinitely at delay 1 and arity 2 deterministically and moreover that the only infinite conformations which its oritatami system can yield is only the zig-zag conformation shown in Fig.\ref{TTT_zigzag}.

@@

Let us consider the stabilization of the $i$-th bead $a_i$ upon $C_{i-1}$. The bead cannot collaborate with any succeeding bead $w[i+1],w[i+2],\cdots$ at delay 1. There are just two ways to get stabilized at delay 1. One way is to be bound to another bead. The other way is through a $tunnel\ section$. A tunnel section consists of four beads that occupy four neighbors of a point (Fig.\ref{TTT_tunnel_intro}). 
How they are stabilized can be described by a binary sequence $S$ of $b$'s (bound) and $t$'s (tunnel section); priority is given to $t$, that is, $S[i] = t$ if the $i$-th bead $w_i$ is stabilized not only by being bound but also through a tunnel section. 
Assume that four of the six neighbors of a point $p$ are occupied by beads $a_{j_1},a_{j_2},a_{j_3},a_{j_4}$ while the other two are free. We call such a bead $p$ as $inside\ of\ a\ tunnel$ and such beads $p^\prime$ as $entrance\ of\ a\ tunnel$ without a case that $p^\prime$ is inside of a tunnel. If the beads $w[i-2]$ and $w[i-1]$ are stabilized respectively at one of the two free neighbors and at $p$ one after another, then the next bead $w[i]$ cannot help but be stabilized at the other free neighbor. In this way, $w[i]$ can get stabilized without being bound.

If a bead is stabilized through a tunnel section, then it can provide some bonds. Let us consider on bonds of $C_i$. $C_i$ is represented $C = (W,P,H)$ where $|W| = i + n$ $C_i$ contains $\alpha \cdot (i + n) - 2|H|$ bonds because $C_i$ consists of $i + n$ beads and a bead has just $\alpha$ bonds and then $2 |H|$ of the those bonds are already used. However, even if a bead has an available bond, $w[j]$ $(j > i)$ might not be able to use this bond because the bond has possibility that it is blocked by transcripts $w[i+1..j-1]$. Number of $binding\ capabilities$ does not contain that case so that it is at most $\alpha \cdot (i + n) - 2|H|$.

@@

\begin{theorem}[Tunnel Troll Theorem]
  Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha \geq 2$. If there is a part of transcript $w[i-1..j+1] = btt...tb$, then $C_i$'s binding capabilities is more than $C_j$'s binding capabilities.
\end{theorem}

\subsection{Proof of Tunnel Troll Theorem}
Assume $\Xi$ is deterministic. %How they are stabilized can be described by a binary sequence $S$ of $b$'s (bound) and $t$'s (tunnel section); priority is given to $t$, that is, $S[i] = t$ if the $i$-th bead $w_i$ is stabilized not only by being bound but also through a tunnel section. 
Each of beads in transcript are bound either inside of tunnel or outside (Fig. \ref{TTT_position1}). If a bead is stabilized at inside of tunnel, then its successor is already decided the position either inside of a tunnel or outside. Moreover, If a bead is stabilized at outside of a tunnel, then its position is an entrance of a tunnel or otherwise.

Tunnel sections have three possible shape with considering symmetry such as straight, acute turn and obtuse turn (Fig. \ref{TTT_tunnel}). Let us focus entrances of a tunnel such as $p^\prime$, then Entrances have two possible shape (Fig. \ref{TTT_enterpoint}).

@@

      \begin{minipage}{0.2\hsize}
      \centering
        \begin{tikzpicture}

@@

      \begin{minipage}{0.2\hsize}
      \centering
        \begin{tikzpicture}

@@

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
	 \draw[->,white] (120:1.1)--(120:0.1);%for adjustment

            \fill(0,0) circle [radius=0.1];
            \node[below] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(60:1.2)},right] {A};
            \node[transform canvas={shift=(-60:1.2)},right] {B};
            \node[transform canvas={shift=(120:1.2)},left] {c};
            \node[transform canvas={shift=(-120:1.2)},left] {d};

          \node at (0,-2) {$t_{0}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}

            \fill(0,0) circle [radius=0.1];
            \node[below] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(180:1.2)},left] {c};
            \node[transform canvas={shift=(-120:1.2)},left] {d};

          \node at (0,-2) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
            \fill(0,0) circle [radius=0.1];
            \node[above] at (0,0) {$p^\prime$};

@@

            \node[transform canvas={shift=(120:1.2)},left] {c};
            \node[transform canvas={shift=(180:1.2)},left] {d};
            \node[transform canvas={shift=(-120:1.2)},left] {e};
            \node[transform canvas={shift=(0:1.2)},below] {n};

          \node at (0,-2) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

@@

  \end{itemize}
\end{itemize}
\begin{figure}
  \centering
    \begin{tabular}{cc}

      \begin{minipage}{0.48\hsize}
      \centering
        \begin{tikzpicture}

@@

          \foreach \theta in {60,-60,120,-120,180}{
            \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.2)}](0.1,-0.1)--(-0.1,0.1);
          }

          \fill(0:1.2) circle [radius=0.1];
          \draw[->] (0:0.1)--(0:1.1);

          \node[left] at (-60:1.2) {A};

          \begin{scope}[shift=(0:2.4)]
            \fill(0,0) circle [radius=0.1];

            \foreach \theta in {120,60,-60}{
              \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1)--(0.1,0.1);
              \draw[transform canvas={shift=(\theta:1.2)}](0.1,-0.1)--(-0.1,0.1);
            }

            \draw[->] (180:1.1)--(180:0.1);
            \draw[->] (-120:0.1)--(-120:1.1);

            \node[below] at (180:1.2) {$w_{i-2}$};
            \node[right] at (0:0) {$w_{i-1}$};
            \node[right] at (-120:1.2) {$w_i$};

            \node[left] at (120:1.2) {B};
            \node[left] at (60:1.2) {C};
            \node[left] at (0:1.2) {D};

            \begin{scope}[shift=(-120:1.2)]
              \draw(0,0) circle [radius=0.1];
              \foreach \theta in {-120,-60}{
                \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1) rectangle (0.1,0.1);
              }

              \node[right] at (-120:1.2) {e};
              \node[right] at (-60:1.2) {f};
            \end{scope}

@@

          \node at (1.2,-2.5) {Pattern 1};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.48\hsize}
      \centering
        \begin{tikzpicture}

@@

          \foreach \theta in {60,0,120,-120,180}{
            \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.2)}](0.1,-0.1)--(-0.1,0.1);
          }

          \draw[->] (-60:0.1)--(-60:1.1);

          \begin{scope}[shift=(-60:1.2),shift=(0:1.2)]
            \fill(0,0) circle [radius=0.1];
            \fill(180:1.2) circle [radius=0.1];

            \foreach \theta in {120,60,0,-60}{
              \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1)--(0.1,0.1);
              \draw[transform canvas={shift=(\theta:1.2)}](0.1,-0.1)--(-0.1,0.1);
            }

            \draw[->] (180:1.1)--(180:0.1);
            \draw[->] (-120:0.1)--(-120:1.1);

            \node[below] at (180:1.2) {$w_{i-2}$};
            \node[right] at (0:0) {$w_{i-1}$};
            \node[right] at (-120:1.2) {$w_i$};

            \node[left] at (120:1.2) {A};
            \node[left] at (60:1.2) {B};
            \node[left] at (0:1.2) {C};

            \begin{scope}[shift=(-120:1.2)]
              \draw(0,0) circle [radius=0.1];
              \foreach \theta in {180,-120,-60}{
                \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1) rectangle (0.1,0.1);
              }

              \node[right] at (180:1.2) {e};
              \node[right] at (-120:1.2) {f};
              \node[right] at (-60:1.2) {g};
            \end{scope}

@@

          \node at (1.25,-4) {Pattern 2};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Case of $S[i-2..i] = tbt$}
    \label{TTT_tunnelC_enter_usingTunnel}
\end{figure}

@@

\begin{figure}
  \begin{center}
        \begin{tikzpicture}
        \draw[->,white] (-70:0.1) -- (-70:3);
          \foreach \theta in {60,-60,120,0}{
            \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1)--(0.1,0.1);
            \draw[transform canvas={shift=(\theta:1.2)}](0.1,-0.1)--(-0.1,0.1);
          }

          \fill(180:1.2) circle [radius=0.1];
          \fill(0:0) circle [radius=0.1];

          \draw[transform canvas={shift=(180:1.2)}] (105:1.1) edge[bend right] (105:0.1);
          \draw[->] (180:1.1) -- (180:0.1);
          \draw[->] (-120:0.1) -- (-120:1.1);

          \node[below] at (180:1.2) {$t_1$};
          \node[right] at (0:0) {$t_2$};
          \node[left] at (-120:1.2) {$t_3$};

          \node[right] at (120:1.2) {A};
          \node[right] at (60:1.2) {B};
          \node[right] at (0:1.2) {C};

          \begin{scope}[shift=(-120:1.2)]
            \draw(0,0) circle [radius=0.1];
            \foreach \theta in {-120,-60,180}{
              \draw[transform canvas={shift=(\theta:1.2)}](-0.1,-0.1) rectangle (0.1,0.1);
            }

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[i-2..i] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \begin{tikzpicture}

      \draw[transform canvas={shift=(-120:1.2)}](-0.1,-0.1)--(0.1,0.1);
      \draw[transform canvas={shift=(-120:1.2)}](0.1,-0.1)--(-0.1,0.1);

      \fill(0:0) circle [radius=0.1];
      \fill(0:1.5) circle [radius=0.1];
      \fill[shift=(0:1.2)](40:1.2) circle [radius=0.1];

      \draw[->] (-120:1.1) edge[bend left] (-120:0.1);
      \draw[->] (0:0) -- (0:1.3);
      \draw[->,transform canvas={shift=(0:1.2)}] (0:0.1) edge[bend right] (40:1.3);

      \draw [shift=(60:1.2)] (-0.1,-0.1) rectangle (0.1,0.1);
      \draw [shift=(-60:1.2)] (-0.1,-0.1) rectangle (0.1,0.1);

      \node[below] at (-120:1.2) {$v_{i-2}$};
      \node[above] at (0:0) {$v_{i-1}$};
      \node[above] at (0:1.2) {$v_i$};
      \node[left, shift=(0:1.2)] at (40:1.2) {$v_{i+1}$};

      \node[right] at (60:1.2) {A};
      \node[right] at (-60:1.2) {B};

    \end{tikzpicture}
    \caption{Case of $S[i]$}
    \label{TTT_case_of_o}
  \end{center}
\end{figure}
